### 2013-09-28 selection_sim_results.tex
Added a few existing files

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Fri Sep 27 17:35:50 2013
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & Fdp & R & S & V & Power \\ 
  \hline
first & 0.06 & 10.48 & 9.77 & 0.71 & 0.98 \\ 
  forward & 0.00 & 0.01 & 0.01 & 0.00 & 0.00 \\ 
  hybrid & 0.08 & 2.00 & 1.84 & 0.16 & 0.18 \\ 
   \hline
\end{tabular}
\caption{Evaluation of model selection using several stopping rules based on our p-values. The naive stopping rule performs well.} 
\end{table}