### 2013-10-02 selection_sim_results.tex
Almost broke group_lasso_knot trying to introduce Sigma

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Tue Oct  1 23:11:01 2013
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & Fdp & R & S & V & Power \\ 
  \hline
first & 0.06 & 10.57 & 9.86 & 0.71 & 0.99 \\ 
  forward & 0.13 & 1.00 & 0.87 & 0.13 & 0.09 \\ 
  hybrid & 0.12 & 2.00 & 1.76 & 0.24 & 0.18 \\ 
   \hline
\end{tabular}
\caption{Evaluation of model selection using several stopping rules based on our p-values. The naive stopping rule performs well.} 
\end{table}