### 2014-05-12 tables/small.tex
Paper edits, adding figs/tables

% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Wed May  7 22:07:35 2014
\begin{table}[ht]
\centering
\begin{tabular}{r|rrrrrrrr}
  \hline
Step & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 \\ 
  \hline
Variable & 1 & 9 & 2 & 8 & 4 & 7 & 3 & 10 \\ 
T$\chi$ p-value & 0.00 & 0.01 & 0.15 & 0.50 & 0.77 & 0.81 & 0.84 & 0.92 \\ 
$\chi^2$ p-value & 0.00 & 0.00 & 0.04 & 0.21 & 0.29 & 0.56 & 0.73 & 0.82 \\ 
   \hline
\end{tabular}
\caption{\em Results of forward stepwise on this example}
\label{tab:example}
\end{table}