### 2014-05-14 tables/small.tex
Paper possibly ready for arxiv

% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Tue May 13 15:41:11 2014
\begin{table}[ht]
\centering
\begin{tabular}{l|rrrrrrrr}
  \hline
Step & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 \\ 
  \hline
Variable & 1 & 9 & 2 & 8 & 4 & 7 & 3 & 10 \\ 
  \hline
  $T\chi$ & 0.00 & 0.01 & 0.16 & 0.60 & 0.72 & 0.81 & 0.84 & 0.92 \\ 
  $\chi^2$ & 0.00 & 0.00 & 0.04 & 0.21 & 0.29 & 0.56 & 0.73 & 0.82 \\ 
  max-$\chi$ & 0.00 & 0.04 & 0.50 & 0.98 & 0.98 & 0.99 & 0.99 & 0.98 \\ 
   \hline
\end{tabular}
\caption{\em Small comparison of $p$-values. Elapsed time for forward stepwise \textit{and} $T\chi$ computation: 0.022 seconds, and for Monte Carlo sample estimate of max-$\chi$ with $M=200$: 0.235 seconds.} 
\label{tab:example}
\end{table}